---
title: "Kaggle_EDA"
output: html_document
---



##a)

Load the training_target.csv file into a data frame in R using the read.csv command (Tip: you can type ?read.csv at the R console to obtain help information about this function). Every row in this data frame corresponds to an ALS (amyotrophic lateral sclerosis) patient in twelve-month clinical trial, identified by a unique subject ID.


```{r}
t_feat = read.csv('../training_features.csv')
t_targ = read.csv('../training_target.csv')
v_feat = read.csv('../validation_features.csv')
v_targ = read.csv('../validation_target.csv')
```


Set na values to the medians of that predictor:
```{r}
t_feat <- as.data.frame(lapply(t_feat, function(x) {x[is.na(x)] <- median(x, na.rm=TRUE); x}))
```

Question: How many patients are in this data frame?
```{r}
dim(t_targ)[1] 
```

```{r}
summary(t_targ)
```


Question: How many feature columns are in this data frame?
```{r}
dim(t_feat)[2]
```

Write code to find the number of NA values in each column, and plot a histogram of these NA counts.

```{r}
num.nas <- function(x) sum(is.na(x))
NAs <- apply(t_feat, 1, num.nas)
hist(NAs, breaks = seq(0, length(NAs), 30))
```


```{r}

```


